Open-source tools in R for landscape ecology

### 2021-01-13 paper/paper.Rmd
updated to include geostats section

@@

<!-- MH: What about spatial autocorrelation? -->
<!-- MH: SpatialKDE, geoR, gstat, spdep, siland -->

>>>>

Geostatistics, or spatial statistics, are complimentary to landscape metrics, and can be used to analyse patterns in continuous data (e.g. normalised difference vegetation index, disturbance intensity, topography). 
In landscape ecology, geostatistics has three key uses: (1) detecting and correcting for spatial autocorrelation; (2) quantifying and comparing landscape pattern; (3) interpolating data.

Point pattern analysis is one such method. 
This uses event-level data, such as locations of individuals, and links the spatial pattern to the ecological process.
The *spatstat* packge contains functionality for point pattern analysis, including exploratory analysis; simulation of point process models; and modelling fitting, inference and diagnostics.

Distance-based methods allow us to detect and correct for spatial autocorrelation in data.
It is key to do so as landscape data are highly spatially autocorrelated, and this non-independence can affect inferences from statistical modelling. 
The *spdep* package has methods for quantifying multiple metrics of spatial autocorrelation, and to correct for these in a spatial autoregressive model.
The *rinla* and *inlabru* packages also provide functionality for modelling of spatially structured data.

Finally, with variograms and correlograms, the spatial structure of continuous landscapes can be quantified and compared. 
R packages *geoR* and *gstat* provide functionality for this type of analysis, as well as interpolation methods, known as (co-)kriging.

@@

### 2021-01-13 paper/paper.tex
updated to include geostats section

>>>>

Geostatistics, or spatial statistics, are complimentary to landscape metrics, and can be used to analyse patterns in continuous data (e.g.~normalised difference vegetation index, disturbance intensity, topography).
In landscape ecology, geostatistics has three key uses: (1) detecting and correcting for spatial autocorrelation; (2) quantifying and comparing landscape pattern; (3) interpolating data.

Point pattern analysis is one such method.
This uses event-level data, such as locations of individuals, and links the spatial pattern to the ecological process.
The \emph{spatstat} packge contains functionality for point pattern analysis, including exploratory analysis; simulation of point process models; and modelling fitting, inference and diagnostics.

Distance-based methods allow us to detect and correct for spatial autocorrelation in data.
It is key to do so as landscape data are highly spatially autocorrelated, and this non-independence can affect inferences from statistical modelling.
The \emph{spdep} package has methods for quantifying multiple metrics of spatial autocorrelation, and to correct for these in a spatial autoregressive model.
The \emph{rinla} and \emph{inlabru} packages also provide functionality for modelling of spatially structured data.

Finally, with variograms and correlograms, the spatial structure of continuous landscapes can be quantified and compared.
R packages \emph{geoR} and \emph{gstat} provide functionality for this type of analysis, as well as interpolation methods, known as (co-)kriging.
